Network Configuration Synthesis with Abstract Topologies \vspace{1cm}

### 2017-03-18 paper/abstraction/paper.tex
edits to abstract, intro, acks, header

\authorinfo{Ryan Beckett$^*$\and Ratul Mahajan$^\dagger$\and Todd Millstein$^\S$\and Jitendra Padhye$^\dagger$\and David Walker$^*$}
{Princeton University$^*$, Microsoft Research$^\dagger$, University of California Los Angeles$^\S$}
{\{rbeckett,dpw\}@cs.princeton.edu, \{ratul,padhye\}@microsoft.com, todd@cs.ucla.edu}

@@

\textbf{Abstract---}
We develop \sysname, a system to synthesize provably-correct
BGP configurations for large, evolving networks from high-level
specifications of topology, routing policy, and fault-tolerance
requirements. \sysname is based on new abstractions for capturing
parameterized network topologies and their evolution, and algorithms
to analyze the impact of topology and routing policy on fault
tolerance. Our algorithms operate entirely on abstract topologies.  We
prove that the properties established by our analyses hold for every
concrete instantiation of the given abstract topology. \sysname
also guarantees that minimal changes to existing device configurations
are required when the network evolves to add or remove devices and
links. Our experiments with real-world topologies and policies show
that our abstractions and algorithms are effective, and that, for large
networks, \sysname synthesizes configurations two orders of magnitude
faster than systems that operate on concrete topologies.

@@

Unfortunately, templates use the same low-level constructs as ordinary router configurations (\EG,
add a tag to this particular announcement; remove a tag from that announcement).  In both templatized
and non-templatized configurations, ensuring that a collection of low-level, local configurations
achieves some important global, network-wide policy, such as a guarantee of connectivity in the face of 
failures, is extremely challenging.
%
%Worse, operators can never be sure that their templates correctly implement the policy. %
%Formally verifying them is difficult; today's hand-generated templates have no particular semantics---which 
%policies they implement and for which topologies---other than those of the configurations obtained from 
%instantiating them.
%
In order to validate their templates, operators will typically first instantiate a template with appropriate
concrete parameters
and then test it using various scenarios.  In general, like in any complex software system, such testing is 
bound to be incomplete.
%Thus, operators validate templates by testing some concrete instantiations using
%simulation or verification tools~\cite{feamster+:rcc, batfish, era, arc, bagpipe}. 
Moreover, even if network operators were to instantiate their templates using their initial network 
configuration and to verify key properties using a tool such as Bagpipe~\cite{bagpipe}, the guarantees
provided would not hold as large networks evolve frequently over time.  In particular,
devices and links are sometimes taken offline for maintenance and sometimes added to expand
capacity.  Unfortunately, templates that
work for the current topology may or may not work for future
topologies. Problems that occur may cause operators to make small, non-uniform changes to particular router configurations,
which defeats the purpose of a template system.  An even worse situation occurs when
operators feel obligated to update many devices to extend or change their network.  
Such network-wide configuration changes entail a great deal of risk and may be highly
disruptive in production networks with live traffic.
%Ideally, network engineers would like some guarantees about the future, namely that 
%certain types of changes to the network topology may be (1) made \emph{locally}, and do not require
%updates of routers far from the source of the change, and (2) do not violate key properties, such
%as fault tolerance.
%Updating the templates, if found to not work for a future topology,
%may cause trickle-down, network-wide configuration changes that are highly
%disruptive in production networks with live traffic.
%Building sophisticated tools for validating templates is difficult

@@

Given the challenges of generating and validating network configurations, one
might think that operators will be receptive to systems that can generate
provably-correct configurations from high-level policy
specifications~\cite{narain+:configassure,propane,synet-synthesis}. However, our conversations
with two major cloud providers reveal that operators of large networks are
reluctant to use such systems. While they think of their network abstractly, in
terms of roles, synthesis systems operate over concrete topologies.  
Even if two devices play the same role, operators cannot specify policy in terms of
this role and even if specifications for the two devices are similar,
there is no guarantee that the systems will generate (syntactically) similar
configurations.  Perhaps most importantly, if the operators want to debug or analyze system output, they
will have to consider hundreds of device configurations instead of just a handful of
role configurations. Finally, current synthesis systems are brittle in the face of network
evolution: Any change in network topology requires re-execution of the engine,
from scratch, on the new topology.  The result may be a completely different
set of configurations.  No administrator of a large network is
able to shut down a production network, upgrade policy on all devices and 
then restart their network.

To address the challenge of configuration synthesis in the presence of abstract roles, we present \sysname. 
%BGP, unlike link state protocols
%such as OSPF, is also particularly
%useful in our setting because it contains features that help us localize the effects of changes.
\sysname
allows operators to input \emph{abstract topologies} in terms of roles
and their connectivity. For instance, they may specify roles for ``top-of-rack''
and ``aggregation" routers and specify that every top-of-rack router connects to
at least two aggregation routers (to tolerate the failure of a link to an
aggregation router). Other inputs to \sysname include the routing policy of the
network and its fault-tolerance requirements (\EG, the number of simultaneous link failures
to tolerate). For high-level specification of routing policy, we borrow notation from the original
\Propane system~\cite{propane}.  However, while \Propane policies refer to concrete devices,
\sysname policies refer to abstract roles.

Based on these inputs, \sysname generates one template per role. 
These templates specify routing policy using the Border Gateway Protocol (BGP).
BGP is a useful target for synthesis, as it is the standard protocol
for interdomain routing and it is also commonly used
within data centers due to its scalability.  However, it is far from the only control
plane protocol; synthesis of configurations for other protocols such as OSPF is beyond the 
scope of this paper.

During synthesis, our compiler performs an analysis over abstract topologies to determine
the fault tolerance properties of the specified policy.  This analysis determines a worst-case 
lower bound on the number of link failures require to disconnect one abstract location from another 
given the constraints specified by the abstract topology. Any concrete instance of the abstract topology will adhere to the given fault tolerance property.
%% first analysis determines whether a user policy can be implemented correctly using BGP in
%% the face of failures (\ie, given any number of failures, will the traffic flow along one of the 
%% specified paths if such a path is available).  The second analysis determines the number of failures
%% required to disconnect one abstract location from another.  
%% Our synthesis algorithm comes with a strong guarantee come with a strong
%% guarantee:  when the templates are instantiated with parameters, giving rise to concrete configurations,
%% and those concrete configurations  {\em
%% any} concrete topology that complies with the abstract topology, the resulting
%% network will correctly implement the routing policy and meet the fault-tolerance
%% requirements.

@@

\sysname provides two other benefits. First, its templates are evolution friendly. 
When the network migrates from
one concrete topology to another, both of which comply with the abstract
topology, only the configurations of devices that acquire or lose a neighbor need to change. 
This guarantee is the best that any system can give as neighbor relationships are explicitly 
configured in devices.  We achieve this guarantee in part by relying on the ability of BGP
to tag announcements with abstract values.  These abstract values make it possible to change
the source of an announcement without the target being aware of a change happening.  Other control
plane protocols that do not provide such tagging mechanisms or that implement link-state protocols as opposed to
path-vector protocols would not be able to avail themselves of these techniques.
%No other device in the network is impacted.

Second, \sysname can scale substantially better than its nearest competitor, the original \Propane system, 
which operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgery}.
Consequently, \sysname can be two orders of magnitude
faster than \Propane, taking less than 10 seconds to synthesize templates for large structured networks.

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, our contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item New topology abstractions for network programming
based on graph homomorphisms and connectivity rules.
\item New algorithms for analyzing routing policy and fault tolerance over
abstract topologies.
%one which guarantees correct compilation and one which determines fault tolerance
%properties of network policy.
%Both algorithms provide guarantees for any
%concrete network satisfying the abstract connectivity invariants.
%\item Design of a template generation strategy that supports network
%evolution.
\item Implementation and evaluation of a \sysname compiler that generates
BGP templates and is much faster
than compilers based on concrete topologies.
\end{itemize}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=2.5in]{figures/example2}
  \vspace{-.8em}
  \caption{A modified version of the example in Figure~\ref{fig:example}.}
  \label{fig:example2}
  \vspace{-.8em}
\end{figure}

@@

@@

\para{Acknowledgments}
We thank the PLDI reviewers for comments on the paper. This work is supported in part by the National Science Foundation award CNS-1161595 as well as Cisco award 677002.

>>>>

\authorinfo{Ryan Beckett$^*$\and Ratul Mahajan$^{\dagger\ddagger}$\and Todd Millstein$^{\dagger\S}$\and Jitendra Padhye$^\ddagger$\and David Walker$^*$}
{Princeton University$^*$ \hspace{0.2in} Intentionet$^\dagger$ \hspace{0.2in}  Microsoft$^\ddagger$ \hspace{0.2in}  University of California Los Angeles$^\S$}
%{\{rbeckett,dpw\}@cs.princeton.edu, \{ratul,padhye\}@microsoft.com, todd@cs.ucla.edu}

@@

\textbf{Abstract---}
We develop \sysname, a system to synthesize provably-correct
BGP (border gateway protocol) configurations for large, evolving networks from high-level
specifications of topology, routing policy, and fault-tolerance.
requirements. \sysname is based on new abstractions for capturing
parameterized network topologies and their evolution, and algorithms
to analyze the impact of topology and routing policy on fault
tolerance. Our algorithms operate entirely on abstract topologies.  We
prove that the properties established by our analyses hold for every
concrete instantiation of the given abstract topology. \sysname
also guarantees that minimal changes to existing device configurations
are required when the network evolves to add or remove devices and
links. Our experiments with real-world topologies and policies show
that our abstractions and algorithms are effective, and that, for large
networks, \sysname synthesizes configurations two orders of magnitude
faster than systems that operate on concrete topologies.

@@

Unfortunately, templates use the same low-level constructs as ordinary router configurations (\EG,
adding or removing tags from particular announcements).  In both templatized
and non-templatized configurations, ensuring that a collection of low-level, local configurations
achieves some important global, network-wide policy, such as a guarantee of connectivity in the face of 
failures, is extremely challenging.
%
%Worse, operators can never be sure that their templates correctly implement the policy. %
%Formally verifying them is difficult; today's hand-generated templates have no particular semantics---which 
%policies they implement and for which topologies---other than those of the configurations obtained from 
%instantiating them.
%
In order to validate  their templates, operators will typically first instantiate a template with appropriate
concrete parameters
and then test it under various scenarios.  In general, like in any complex software system, such testing is inherently incomplete.
%Thus, operators validate templates by testing some concrete instantiations using
%simulation or verification tools~\cite{feamster+:rcc, batfish, era, arc, bagpipe}.

Moreover, even if network operators were to instantiate their templates using the initial network 
topology and verify key properties using tools such as Bagpipe~\cite{bagpipe}, the guarantees
would not hold as the network topology evolves. Evolution of the topology is a frequent event for large networks, as  devices and links are taken offline for maintenance and  added to expand
capacity.  
Templates that
work for the current topology may or may not work for future
topologies. Ensuing problems may cause operators to make non-uniform changes to particular routers' configurations,
which defeats the purpose of a template system.  An even worse situation is when
operators have to update many devices to evolve their network.  
Such network-wide configuration changes entail a great deal of risk and can be highly
disruptive to live traffic.
%Ideally, network engineers would like some guarantees about the future, namely that 
%certain types of changes to the network topology may be (1) made \emph{locally}, and do not require
%updates of routers far from the source of the change, and (2) do not violate key properties, such
%as fault tolerance.
%Updating the templates, if found to not work for a future topology,
%may cause trickle-down, network-wide configuration changes that are highly
%disruptive in production networks with live traffic.
%Building sophisticated tools for validating templates is difficult

@@

Given the challenges of generating and validating network configurations, one
might think that operators will be receptive to systems that can generate
provably-correct configurations from high-level policy
specifications~\cite{narain+:configassure,propane,synet-synthesis}. However, our conversations
with two major cloud providers reveal that operators of large networks are
reluctant to use such systems. While they think of their network abstractly, in
terms of roles, current synthesis systems operate over concrete topologies.  
Even if two devices play the same role, operators cannot specify policy in terms of
this role; and even if specifications for the two devices are similar,
there is no guarantee that the systems will generate (syntactically) similar
configurations.  Perhaps most importantly, if the operators want to debug or analyze system output, they
will have to consider hundreds of device configurations instead of just a handful of
role configurations. Current synthesis systems are also brittle in the face of network
evolution. Any change in network topology requires re-execution of the engine,
from scratch, on the new topology, and the result may be a completely different
set of configurations.  No operator is
able to shut down a large, production network, upgrade policy on all devices and 
then restart their network.

To address the challenge of configuration synthesis in the presence of abstract roles, we develop \sysname. 
%BGP, unlike link state protocols
%such as OSPF, is also particularly
%useful in our setting because it contains features that help us localize the effects of changes.
\sysname
allows operators to input \emph{abstract topologies} in terms of roles
and their connectivity. For instance, they may specify roles for ``top-of-rack''
and ``aggregation" routers and specify that every top-of-rack router connects to
at least two aggregation routers (to tolerate the failure of a link to an
aggregation router). \sysname takes two additional inputs. 
The first is a  high-level specification of routing policy, for which we borrow notation from the original
\Propane system~\cite{propane}.  While \Propane policies refer to concrete devices,
\sysname policies refer to abstract roles.
The final input to \sysname is the fault-tolerance requirements of the network, such as the number of simultaneous link failures it can tolerate without loss of connectivity for any traffic flow.

Based on these inputs, \sysname generates one template per role. 
These templates specify routing policy using BGP (border gateway protocol). 
BGP is the standard protocol for interdomain routing and is also commonly used within data centers because of its scalability and support for rich policies.\footnote{However, BGP is not the only routing protocol used in existing networks. In the future, we will extend our work to  other protocols such as OSPF.}
%
Our templates are correct by design and evolution friendly. 
When the network evolves from
one concrete topology to another, both of which comply with the abstract
topology, only the configurations of devices that acquire or lose a neighbor need to change. 
This guarantee is the best that any system can give as neighbor relationships are explicitly 
configured in devices.  We achieve it in part by leveraging BGP features that allow arbitrary tags on routing announcements and expressing policy using such tags (instead of router identifiers). 
%These features makes it possible to change network topologies without reflecting those changes in the configurations of distant routers.  
%Other control plane protocols that do not provide such tagging mechanisms or that implement link-state protocols as opposed to path-vector protocols would not be able to avail themselves of these techniques.
%No other device in the network is impacted.

During synthesis, our compiler analyzes abstract topologies to determine
the fault tolerance properties of the specified routing policy.  This analysis yields a lower  
bound on the number of link failures require to disconnect one abstract location from another. 
%given the connectivity constraints of the abstract topology. 
{\em Any} concrete instance of the abstract topology will adhere to the given fault tolerance property.

%% first analysis determines whether a user policy can be implemented correctly using BGP in
%% the face of failures (\ie, given any number of failures, will the traffic flow along one of the 
%% specified paths if such a path is available).  The second analysis determines the number of failures
%% required to disconnect one abstract location from another.  
%% Our synthesis algorithm comes with a strong guarantee come with a strong
%% guarantee:  when the templates are instantiated with parameters, giving rise to concrete configurations,
%% and those concrete configurations  {\em
%% any} concrete topology that complies with the abstract topology, the resulting
%% network will correctly implement the routing policy and meet the fault-tolerance
%% requirements.

@@

%\sysname provides two other benefits. First, its templates are evolution friendly. 
%When the network migrates from
%one concrete topology to another, both of which comply with the abstract
%topology, only the configurations of devices that acquire or lose a neighbor need to change. 
%This guarantee is the best that any system can give as neighbor relationships are explicitly 
%configured in devices.  We achieve this guarantee in part by relying on the ability of BGP
%to tag announcements with abstract values.  These abstract values make it possible to change
%the source of an announcement without the target being aware of a change happening.  Other control
%plane protocols that do not provide such tagging mechanisms or that implement link-state protocols as opposed to
%path-vector protocols would not be able to avail themselves of these techniques.
%%No other device in the network is impacted.

We use \sysname for a range of real-world topologies and routing policies. We find that it can effectively encode the topologies and their evolution, and the fault tolerance bounds that it computes are often precise. 
We also find that \sysname scales substantially better than its nearest competitor, the original \Propane system, 
which operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgery}.
Consequently, \sysname can be two orders of magnitude
faster than \Propane, taking less than 10 seconds to synthesize templates for large  networks.

\vspace{0.5em}
\noindent
\textbf{Contributions}: To summarize, our contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item New topology abstractions for network programming
based on graph homomorphisms and connectivity rules.
\item New algorithms for analyzing routing policy and fault tolerance over
abstract topologies.
%one which guarantees correct compilation and one which determines fault tolerance
%properties of network policy.
%Both algorithms provide guarantees for any
%concrete network satisfying the abstract connectivity invariants.
%\item Design of a template generation strategy that supports network
%evolution.
\item Implementation and evaluation of a \sysname compiler that generates
BGP templates and is much faster
than compilers based on concrete topologies.
\end{itemize}

@@

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=2.5in]{figures/example2}
  \vspace{-.8em}
  \caption{A modified version of the network in Figure~\ref{fig:example}.}
  \label{fig:example2}
  \vspace{-.8em}
\end{figure}

@@

\para{Acknowledgments}
We thank the many network engineers, especially George Chen and Lihua Yuan, who helped us understand the importance of working with abstract topologies and synthesizing templates. 
We also thank the PLDI reviewers for comments on the paper. This work is supported in part by the National Science Foundation award CNS-1161595 as well as Cisco award 677002.
